### 2024-04-16 chap_RobPol/chap_RobPol.tex
RobPol chapter: submission for review

\subsection{Principal components of the transition}
\label{subsec:RobPol:PC_transition}
Based on the \gls{PCs} of each representative year, $\text{PC}_{y}$, the \gls{PCs} of the transition, \ie the metrics to assess the robustness of roadmaps, can be computed. Before aggregating and averaging similar $\text{PC}_{y}$, it is necessary to rank them to ensure capturing most of the transition variance in the subsequent $\text{PC}_{\text{transition}}$.

@@

\section{Robustness assessment of pathway roadmaps}
\label{sec:RobPol:Rob_Assessment}
Now that the performance metric is defined, we can assess the robustness of different roadmaps. These roadmaps are defined as the technological mix given by the deterministic optimisation of the perfect foresight pathway under certain conditions. The two first roadmaps are the \textbf{REF} and \textbf{SMR} cases (see Chapter \ref{chap:atom_mol}). All the uncertain parameters are considered at their nominal value in these two cases. The only difference consists in allowing \gls{SMR} from 2040 onward in the \textbf{SMR} case. On top of these two cases, we add a so-called ``robust'' case, \textbf{ROB}, in the same sense as in the works of \citet{bertsimas2004price} or \citet{Moret2017PhDThesis} accounting for a ``protection parameter''. Then, we assess the robustness of these three roadmaps via a similar ``three-step rolling horizon approach'' as in the work of \citet{moret2020overcapacity}: (i) setting the initial investment strategies provided by the roadmaps and, (ii) assessing the variation of investments needed through myopic pathway optimisation under uncertainties and, (iii) projecting the strategies from the myopic optimisations on the robustness metrics defined via \gls{PCA}.

\subsection{A robust roadmap?}
\label{subsec:RobPol:Rob_roadmap}
Rather than considering all the uncertain parameters at their worst values as in the work of \citet{soyster1973convex}, we follow here the robust approach of \citet{bertsimas2004price} and \citet{Moret2017PhDThesis}. In their works, the authors consider a factor $\Gamma_{\text{obj}}\in [0,d]$ that represents a ``protection parameter'' of the objective function where $d$ is the total number of uncertain parameters (see Section \ref{subsec:pce}). If $\Gamma_{\text{obj}}= 0$, where all the uncertain parameters are at their nominal value, we obtain the deterministic solution of the \textbf{REF} case. If $\Gamma_{\text{obj}}= d$, this means considering the ``fully robust'' solution where all the uncertain parameters are their worst value, as in \citet{soyster1973convex}.  Between these two extreme cases, the uncertain parameters to account for at their worst values follow the ranking given by the \gls{GSA} (see Chapter \ref{chap:atom_mol}). In practice, $\Gamma_{\text{obj}}= 1$ means considering the cost of purchasing electrofuels at its worst value. $\Gamma_{\text{obj}}= 2$ adds the industry \gls{EUD} at its worst value in the lot, and so on. In the present analysis, we consider the ``robust'' roadmap by setting $\Gamma_{\text{obj}}= 6$. Consequently, costs of purchasing electrofuels, fossil fuels and biofuels as well as the industry \gls{EUD}, the interest rate and the variable OPEX of technologies are considered at their worst values, \ie the upper bound of their respective range. In practice, this means that, in the ROB case, imported energy carriers, except electricity, are 179.8\% more expensive, the industrial demand is, by 2050, 23.9\% higher\footnote{As detailed in Chapter \ref{chap:case_study}, the industrial demand encompasses the whole \gls{NED} and \gls{LT} heat demand, as well as the industrial share of electricity and \gls{LT} heat demands.}, the interest rate is 46.2\% (\ie 2.2\% versus 1.5\% in the REF case) and the variable OPEX of technologies, $c_{\text{\emph{maint}}}$, is 35.7\% higher.  The following paragraphs investigate the differences in the different sectors between the ROB and the REF cases, similarly to the comparison carried out with the SMR case in Chapter \ref{chap:atom_mol}.\\

@@

\section{Discussion}
\label{sec:RobPol:Discussion}

>>>>

\subsection{Principal components of the perfect foresight REF transitions}
\label{subsec:RobPol:PC_transition}
Based on the \gls{PCs} of each representative year, $\text{PC}_{y}$, the \gls{PCs} of the transition, \ie the metrics to assess the robustness of roadmaps, can be computed. Before aggregating and averaging similar $\text{PC}_{y}$, it is necessary to rank them to ensure capturing most of the transition variance in the subsequent $\text{PC}_{\text{transition}}$.

@@

\newpage
\section{Robustness assessment of pathway roadmaps}
\label{sec:RobPol:Rob_Assessment}
Now that the performance metric is defined, we can assess the robustness of different roadmaps. These roadmaps are defined as the technological mix given by the deterministic optimisation of the perfect foresight pathway under certain conditions. The two first roadmaps are the \textbf{REF} and \textbf{SMR} cases (see Chapter \ref{chap:atom_mol}). All the uncertain parameters are considered at their nominal value in these two cases. The only difference consists in allowing \gls{SMR} from 2040 onward in the \textbf{SMR} case. On top of these two cases, we add a so-called ``robust'' case, \textbf{ROB}, in the same sense as in the works of \citet{bertsimas2004price} or \citet{Moret2017PhDThesis} accounting for a ``protection parameter''. Then, we assess the robustness of these three roadmaps via a similar ``three-step rolling horizon approach'' as in the work of \citet{moret2020overcapacity}: (i) setting the initial investment strategies provided by the roadmaps and, (ii) assessing the variation of investments needed through myopic pathway optimisation under uncertainties and, (iii) projecting the strategies from the myopic optimisations on the robustness metrics defined via \gls{PCA}.

\subsection{A robust roadmap?}
\label{subsec:RobPol:Rob_roadmap}
Rather than considering all the uncertain parameters at their worst values as in the work of \citet{soyster1973convex}, we follow here the robust approach of \citet{bertsimas2004price} and \citet{Moret2017PhDThesis}. In their works, the authors consider a factor $\Gamma_{\text{obj}}\in [0,d]$ that represents a ``protection parameter'' of the objective function where $d$ is the total number of uncertain parameters (see Section \ref{subsec:pce}). If $\Gamma_{\text{obj}}= 0$, where all the uncertain parameters are at their nominal value, we obtain the deterministic solution of the \textbf{REF} case. If $\Gamma_{\text{obj}}= d$, this means considering the ``fully robust'' solution where all the uncertain parameters are their worst value, as in \citet{soyster1973convex}.  Between these two extreme cases, the uncertain parameters to account for at their worst values follow the ranking given by the \gls{GSA} (see Chapter \ref{chap:atom_mol}). In practice, $\Gamma_{\text{obj}}= 1$ means considering the cost of purchasing electrofuels at its worst value. $\Gamma_{\text{obj}}= 2$ adds the industry \gls{EUD} at its worst value in the lot, and so on. In the present analysis, we consider the ``robust'' roadmap by setting $\Gamma_{\text{obj}}= 6$. Consequently, costs of purchasing electrofuels, fossil fuels and biofuels as well as the industry \gls{EUD}, the interest rate and the variable OPEX of technologies are considered at their worst values, \ie the upper bound of their respective range. In practice, this means that, in the ROB case, imported energy carriers, except electricity, are 179.8\% more expensive, the industrial demand is, by 2050, 23.9\% higher\footnote{As detailed in Chapter \ref{chap:case_study}, the industrial demand encompasses the whole \gls{NED} and \gls{LT} heat demand, as well as the industrial share of electricity and \gls{LT} heat demands.}, the interest rate is increased by 46.2\% (\ie 2.2\% versus 1.5\% in the REF case) and the variable OPEX of technologies, $c_{\text{\emph{maint}}}$, is 35.7\% higher.

Besides the additional operational costs due to more expensive energy carriers, this case presents cumulative investments over the transition that are 2.6\% (17\,b€) higher than the REF case. The following paragraphs investigate the differences in the different sectors between the ROB and the REF cases, similarly to the comparison carried out with the SMR case in Chapter \ref{chap:atom_mol}.\\

@@

\subsection{Principal components of myopic transitions}
\label{subsec:RobPol:PC_MY_transition}
When computing the \gls{PCs} of the different myopic transitions, the ROB case has a variance that is 10\% smaller than the one of the REF case. Moreover, we observe that the variations of \gls{PV}, industrial resistors and \gls{LT}-heat technologies still have a major contribution on the variation of the system roadmaps (see Figure \ref{fig:Top5_PCtransition_cases}). More surprising is the contribution of the variation of the technology to produce methanol from methane. Where methanol is preferably imported from abroad (or produced from biomass) in the perfect foresight approach, the myopic optimisation opts for this technology in the near-term (2025) to compete with the imports as the main supplier of this energy carrier.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.49\textwidth]{Top5_PCtransition_REF_MY.pdf}
\includegraphics[width=0.49\textwidth]{Top5_PCtransition_SMR_MY.pdf}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.49\textwidth]{Top5_PCtransition_ROB_MY.pdf}
\caption{Contribution of the top-5 technologies to the \gls{PCs} of myopic transitions for the REF, SMR and ROB cases. The variation of \gls{PV}, industrial electric resistors and \gls{LT}-heat technologies still have a major contribution of the variation in the system roadmaps. At the early stages of the myopic transitions, the installation methane-to-methanol is also subject to significant variations.}
\label{fig:Top5_PCtransition_cases}
\end{figure}

\section{Conclusions}
\label{sec:RobPol:Conclusions}
In this chapter, we applied the \gls{PCA}-based approach detailed in Chapter \ref{chap:chap_methodo} to assess the robustness of different transition roadmaps for the case of Belgium. First, starting from the principal components of each representative year, we have computed the PCs of the transition to serve as robustness metric. This showed that the variation of installed capacities of \gls{PV}, industrial electric resistors and \gls{LT}-heat technologies drove the major share (75\%) of the variation of the transition design strategy.

Then, we have tested three roadmaps resulting from different perfect foresight optimisations: REF, SMR and ROB. Where the first two have been broadly discussed in Chapter \ref{chap:atom_mol}, the ``robust'' (ROB) case was considering the worst value of the uncertain parameters impacting the most the total transition cost following the same approach as \citet{Moret2017PhDThesis}. From these roadmaps setting the minimal installed capacities over the transition, we have assessed the variation the additional capacities needed in myopic optimisations subject to uncertainties. This showed that counting on the installation of \gls{SMR} by 2040 was not providing more or less robustness than the reference case. However, the robust (ROB) investment roadmap --- which, in our case, anticipates the full deployment of local \gls{VRES} and the integration of efficient technologies --- is more robust than the reference scenario. Even though this roadmap represents 2.6\% more cumulative \gls{CAPEX} over the transition than the REF case, it provides more robustness towards the variations of \gls{PV} (12\%) and technologies in the transport of freight (17\%). Overall, this \gls{PCA}-based approach allows evaluating the variation of multiple-year transition pathway of a whole-energy system and assessing its robustness. Doing so, it brings more information than when considering only the variation of the objective function.

### 2024-04-16 chap_atom_mol/chap_atom_mol.tex
RobPol chapter: submission for review

\section{Discussion}
\label{sec:atom_mol:discuss}
Given its lower \gls{LCOE} (40\,€/MWh$_{\text{elec}}$) and the important share of the investment recovered via its salvage value (between 79\% and 96\%), \gls{SMR} is installed as soon as available. It directly substitutes other flexible power generation units (\ie CCGT) and provides, by 2050, 44.6\,TWh, 25\% of the total electricity production. Consequently, this reduces the need to produce electricity via \gls{CHP} and allows increasing by 48\% the high-temperature heat produced by electric heaters. Besides these two sectors, the others (\ie low-temperature heat, mobility and non-energy demand) are marginally or not impacted by the integration of \gls{SMR}.

>>>>

\section{Conclusions}
\label{sec:atom_mol:conclusions}
Given its lower \gls{LCOE} (40\,€/MWh$_{\text{elec}}$) and the important share of the investment recovered via its salvage value (between 79\% and 96\%), \gls{SMR} is installed as soon as available. It directly substitutes other flexible power generation units (\ie CCGT) and provides, by 2050, 44.6\,TWh, 25\% of the total electricity production. Consequently, this reduces the need to produce electricity via \gls{CHP} and allows increasing by 48\% the high-temperature heat produced by electric heaters. Besides these two sectors, the others (\ie low-temperature heat, mobility and non-energy demand) are marginally or not impacted by the integration of \gls{SMR}.

### 2024-04-16 chap_case_study/chap_case_study.tex
RobPol chapter: submission for review

\section*{Conclusion}
\label{sec:cs:conclusion}
Mainly based on data collected during previous works \cite{Moret2017PhDThesis,limpens2021generating}, this thesis has developed the case study along different axes. First, the non-energy demand, that currently represents about 20\% of the final energy consumed in Belgium, has been brought to a level of details similar to the other energy sectors of the system. Previously considered as a fixed demand supplied by fossil gas, it has been detailed to account for final-use products (\ie \acrfull{HVC}, ammonia and methanol) and their respective production routes. Second, electrofuels are now more detailed in the model as well as their different supply and consumption routes. Third, instead of defining \textit{a priori} the \ce{CO2}-trajectory through the transition, this work sets a limit on the overall \ce{CO2}-budget. Then, based on the methodology of \citet{Moret2017}, uncertainty ranges are defined for the 5-year step transition from 2020 to 2050. Finally, on top of adding the possibility to install small modular reactors (SMR), this thesis accounts for updated values for the \gls{EUD} as well as adjusted assumptions regarding the two key drivers of the transition in the private mobility, \ie battery electric vehicles (BEV) and \acrfull{FC} cars.

>>>>

\section{Conclusions}
\label{sec:cs:conclusions}
Mainly based on data collected during previous works \cite{Moret2017PhDThesis,limpens2021generating}, this thesis has developed the case study along different axes. First, the non-energy demand, that currently represents about 20\% of the final energy consumed in Belgium, has been brought to a level of details similar to the other energy sectors of the system. Previously considered as a fixed demand supplied by fossil gas, it has been detailed to account for final-use products (\ie \acrfull{HVC}, ammonia and methanol) and their respective production routes. Second, electrofuels are now more detailed in the model as well as their different supply and consumption routes. Third, instead of defining \textit{a priori} the \ce{CO2}-trajectory through the transition, this work sets a limit on the overall \ce{CO2}-budget. Then, based on the methodology of \citet{Moret2017}, uncertainty ranges are defined for the 5-year step transition from 2020 to 2050. Finally, on top of adding the possibility to install small modular reactors (SMR), this thesis accounts for updated values for the \gls{EUD} as well as adjusted assumptions regarding the two key drivers of the transition in the private mobility, \ie battery electric vehicles (BEV) and \acrfull{FC} cars.
